### 2016-12-28 MSSM/UncOv.tex
Adding more MSSM selection

\section{Overview of uncertainties}
\label{sec:mssm_uncs_overview}
The systematics that are applied are summarised in
table \ref{tab:SystematicUncertainties_mt} for the $\mu\tau_h$
channel, in table \ref{tab:SystematicUncertainties_et} for the $e\tau_h$ channel,
in table \ref{tab:SystematicUncertainties_tt} for the $\tau_h\tau_h$ channel and
in table \ref{tab:SystematicUncertainties_em} for the $e\mu$ channel. Additionally,
the systematic uncertainties for the $Z\rightarrow\mumu$ control region are shown
in table \ref{tab:SystematicUncertainties_zmm}.

\begin{table}[!h]
\begin{center}
\caption{\footnotesize Systematic uncertainties that affect the estimated number of signal
or background events in the $\mu \tau_{h}$ channel. Some of the uncertainties
are different for different processes, where this is the case ranges are given.
Uncertainties marked "Fully correlated" are correlated between the different
decay channels and event categories, those marked "Fully uncorrelated" are
uncorrelated between the different decay channels and event categories.}
 {\scriptsize
 \begin{tabular}{p{2cm}|p{1cm}p{1cm}p{1cm}p{1cm}|p{1cm}p{1cm}p{1cm}p{1cm}|p{3cm}}
     & \multicolumn{8}{|c}{Event yield uncertainty by event category} &  \\
    \hline
\textbf{ Process }
    & \multicolumn{4}{|c}{\textbf{No b-tag}} & \multicolumn{4}{|c}{\textbf{B-tag}} & \textbf{Correlation}           \\
     & SR & OS high $m_{T}$ & SS low $m_{T}$ & SS high $m_{T}$ & SR & OS high $m_{T}$ & SS low $m_{T}$ & SS high $m_{T}$ & \\
    \hline
    \multicolumn{10}{l}{\textbf{Integrated luminosity 13}\TeV} \\
    All from MC    & 6.2\%  &6.2\%  & 6.2\%    & 6.2\% & 6.2\% & 6.2\% & 6.2\% & 6.2\% & Fully correlated\\
    \hline
    \multicolumn{10}{l}{\textbf{Jet energy scale }}\\
    All from MC & 0--1\% & 0--2\% & 0--1\% & 0--1\%& 0--12\% & 3--11\% & 0--15\% & 0--14\% & Fully correlated \\
    \hline
    \multicolumn{10}{l}{\MET \textbf{scale} }\\
    ZTT,W     & 2\% & & 2\% & &2\% &  & 2\% & & Corr. between chn/cat; TTT,TTJ, VVT,VVJ                         \\
    TTT,TTJ,VVT,VVJ & 2\% & & 2\% & &2\% & & 2\% & & uncorr. from ZTT,W \\
    \hline
    \multicolumn{4}{l}{\MET \textbf{resolution}} \\
    ZTT,W     & 2\% & & 2\% & &2\% &  & 2\% & & Corr. between chn/cat; TTT,TTJ,VVT,VVJ                         \\
    TTT,TTJ,VVT,VVJ & 2\% & & 2\% & &2\% & & 2\% & & uncorr. from ZTT,W \\
    \hline
    \multicolumn{10}{l}{\textbf{Muon identification and trigger} }\\
    All from MC & 2\% & 2\% &2\% &2\%  & 2\%  & 2\% & 2\% & 2\% & Fully correlated      \\
    \hline
    \multicolumn{10}{l}{\textbf{Tau-lepton identification}}\\
    ZTT,TTT,VVT, signal     & 5\% & 5\% & 5\% & 5\%   & 5\%  & 5\% & 5\% & 5\% & Fully correlated \\
    ZTT,TTT,VVT, signal     & 3\% & 3\% & 3\% & 3\%   & 3\%  & 3\% & 3\% & 3\% & Corr. between cats, uncorr. between chns    \\
    \hline
    \multicolumn{10}{l}{\textbf{B-tagging efficiency} }\\
    All from MC & 0--3\% & 0--3\% & 0--3\% & 0--2\% & 1--3\% &  0--3\%& 0--3\% & 0--3\%& Fully correlated\\
    W &  &  &  &  & 4\% & 4\% & 6\% & 6\%& Fully correlated\\
    \hline
    \multicolumn{10}{l}{\textbf{B-mistag rate } }\\
    All from MC & & & & & 1--7\% & & & & Fully correlated\\
    \hline
    \multicolumn{10}{l}{\textbf{Tau-lepton energy scale}}\\
    ZTT,TTT,VVT, signal     & shape & shape & shape & shape  & shape & shape & shape & shape & Corr. between cats, uncorr. between chns   \\
    \hline
    \multicolumn{10}{l}{\textbf{High} $P_{T}$ $\tau_h$\textbf{ ID efficiency } } \\
    ZTT,TTT,VVT, signal    & shape & shape & shape & shape  & shape & shape & shape & shape & Corr. between cats, uncorr. between chns   \\
    \hline
    \multicolumn{10}{l}{\textbf{Top quark} $P_{T}$ \textbf{reweighting} }\\
    TTT,TTJ  & shape & shape & shape & shape & shape & shape & shape & shape & Fully correlated    \\
    \hline
    \multicolumn{10}{l}{\textbf{DY reweighting } }\\
    ZTT       & shape & shape & shape & shape  & shape & shape & shape & shape &Fully correlated              \\
    \hline
    \multicolumn{10}{l}{\textbf{Jet}$\rightarrow\tau_h$ \textbf{fake rate reweighting } }\\
    W         & shape &  &  &   & shape &  &  &  &Fully correlated              \\
    \hline
    \multicolumn{10}{l}{\textbf{Normalisation, }$\PZ$ \textbf{production} }\\
    ZL,ZJ       & 4\% & 4\% & 4\% & 4\% & 4\%  & 4\% & 4\% & 4\% & Fully correlated              \\
    ZTT         &  & 4\% & 4\% & 4\% &  & 4\% & 4\% & 4\% & Fully correlated              \\
   \multicolumn{10}{l}{ $Z\to\Pgt\Pgt$\textbf{: $Z\to\Pgt\Pgt / Z\to\Pgm\Pgm$ extrapolation } } \\
    ZTT         & 3\% & 3\% & 3\% & 3\% & 5\% & 5\% & 5\% & 5\% & Fully uncorrelated           \\
    \multicolumn{10}{l}{\textbf{Normalisation,} $\cPqt\cPaqt$}\\
    TTT,TTJ     & 6\% & 6\% & 6\% & 6\%   & 6\% & 6\% &6\% & 6\%  & Fully correlated                \\
    \multicolumn{10}{l}{\textbf{Normalisation, di-boson} } \\
    VVT,VVJ     & 5\% & 5\% & 5\% & 5\%   & 5\% & 5\% & 5\% & 5\% & Fully correlated         \\
   \multicolumn{10}{l}{ \textbf{Normalisation,} $\PZ\to\Pgm\Pgm$: $\Pgm$ \textbf{misidentified as} $\tau_{h}$ }\\
    ZL     & 30\%  & 30\% & 30\% & 30\%   & 30\% & 30\% & 30\% &30\% &Corr. between cats, uncorr. between chns        \\
    \multicolumn{10}{l}{\textbf{Normalisation,} $\PZ$\textbf{+jets : jet misidentified as} $\tau_{h}$ } \\
    ZJ,TTJ     & 20\%  & 20\% &20\% &20\%      & 20\% & 20\% &20\% &20\%  & Corr. between cats, uncorr. between chns  \\
    \hline
    \multicolumn{10}{l}{\textbf{W OS/SS ratio (stat) } } \\
    W & 2\% & 2\% & & &11\% &11\% & & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W OS/SS ratio (syst) } }\\
    W & 8\% & 8\% & & &10\% &10\% & & &Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W low/high} $m_{T}$ \textbf{ratio (stat)}}\\
    W & 2\% & & 2 \% & &14\% & & 14\% & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W low/high} $m_{T}$ \textbf{ratio (syst)} }\\
    W & 20\% & &20\% & & 20\% & & 20\% & & Fully uncorrelated \\
    %\multicolumn{10}{l}{\textbf{W b-tag extrap (stat)}}\\
    %W &  & &  & &12\% & 8\%& 22\% & 14\%& Fully uncorrelated \\
    %\multicolumn{10}{l}{\textbf{W b-tag extrap (syst)} }\\
    %W &  & & & & 10\% & & 10\% & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{QCD OS/SS ratio (syst) }}\\
    QCD & 4\% & 4\% & & & 60\% & 60\% & & &Fully uncorrelated \\
    \hline
\end{tabular}
}
\label{tab:SystematicUncertainties_mt}
\end{center}
\end{table}

\begin{table}[!h]
\begin{center}
\caption{\footnotesize Systematic uncertainties that affect the estimated number of signal
or background events in the $e \tau_{h}$ channel. Some of the uncertainties
are different for different processes, where this is the case ranges are given.
Uncertainties marked "Fully correlated" are correlated between the different
decay channels and event categories, those marked "Fully uncorrelated" are
uncorrelated between the different decay channels and event categories.}
 {\scriptsize
 \begin{tabular}{p{2cm}|p{1cm}p{1cm}p{1cm}p{1cm}|p{1cm}p{1cm}p{1cm}p{1cm}|p{3cm}}
     & \multicolumn{8}{|c}{Event yield uncertainty by event category} &  \\
    \hline
\textbf{ Process }
    & \multicolumn{4}{|c}{\textbf{No b-tag}} & \multicolumn{4}{|c}{\textbf{B-tag}} & \textbf{Correlation}           \\
     & SR & OS high $m_{T}$ & SS low $m_{T}$ & SS high $m_{T}$ & SR & OS high $m_{T}$ & SS low $m_{T}$ & SS high $m_{T}$ & \\
    \hline
    \multicolumn{10}{l}{\textbf{Integrated luminosity 13}\TeV} \\
    All from MC    & 6.2\%  & 6.2\%  & 6.2\%    & 6.2\% & 6.2\% & 6.2\% & 6.2\% & 6.2\% & Fully correlated\\
    \hline
    \multicolumn{10}{l}{\textbf{Jet energy scale }}\\
    All from MC & 0--1\% & 0--1\% & 0--2\% & 0--1\%& 1--14\% & 0--12\% & 0--15\% & 0--15\% & Fully correlated \\
    \hline
    \multicolumn{10}{l}{\MET \textbf{scale}}\\
    ZTT,W     & 2\% & & 2\% & &2\% &  & 2\% & & Corr. between chn/cat, TTT,TTJ,VVT,VVJ                         \\
    TTT,TTJ,VVT,VVJ & 2\% & & 2\% & &2\% & & 2\% & & uncorr. from ZTT,W \\
    \hline
    \multicolumn{10}{l}{\MET \textbf{resolution}} \\
    ZTT,W     & 2\% & & 2\% & &2\% &  & 2\% & & Corr. between chn/cat, TTT,TTJ,VVT,VVJ          \\
    TTT,TTJ,VVT,VVJ & 2\% & & 2\% & &2\% & & 2\% & & uncorr. from ZTT,W \\
    \hline
    \multicolumn{10}{l}{\textbf{Electron identification and trigger} }\\
    All from MC & 2\% & 2\% &2\% &2\%  & 2\%  & 2\% & 2\% & 2\% & Fully correlated      \\
    \hline
    \multicolumn{10}{l}{\textbf{Tau-lepton identification}}\\
    ZTT,TTT,VVT, signal     & 5\% & 5\% & 5\% & 5\%   & 5\%  & 5\% & 5\% & 5\% & Fully correlated \\
    ZTT,TTT,VVT, signal     & 3\% & 3\% & 3\% & 3\%   & 3\%  & 3\% & 3\% & 3\% & Corr. between cats, uncorr. between chns    \\
    \hline
    \multicolumn{10}{l}{\textbf{B-tagging efficiency} }\\
    All from MC & 0--3\% & 0--3\% & 0--4\% & 0--4\% & 1--4\% & 0--2\% & 0--2\% & 0--2\% & Fully correlated\\
    W &  &  &  &  & 5\% & 2\% & 10\% & 3\% & Fully correlated\\
    \hline
    \multicolumn{10}{l}{\textbf{B-mistag rate } }\\
    All from MC & & & & & 0--3\% & & & & Fully correlated\\
    \hline
    \multicolumn{10}{l}{\textbf{Tau-lepton energy scale}}\\
    ZTT,TTT,VVT, signal     & shape & shape & shape & shape  & shape & shape & shape & shape & Corr. between cats, uncorr. between chns   \\
    \hline
    \multicolumn{10}{l}{\textbf{High} $P_{T}$ $\tau_h$\textbf{ ID efficiency } } \\
    ZTT,TTT,VVT, signal    & shape & shape & shape & shape  & shape & shape & shape & shape & Corr. between cats, uncorr. between chns   \\
    \hline
    \multicolumn{10}{l}{\textbf{Top quark} $P_{T}$ \textbf{reweighting} }\\
    TTT,TTJ  & shape & shape & shape & shape & shape & shape & shape & shape & Fully correlated    \\
    \hline
    \multicolumn{10}{l}{\textbf{DY reweighting } }\\
    ZTT       & shape & shape & shape & shape  & shape & shape & shape & shape &Fully correlated              \\
    \hline
    \multicolumn{10}{l}{\textbf{Jet}$\rightarrow\tau_h$ \textbf{fake rate reweighting } }\\
    W         & shape &  &  &   & shape &  &  &  &Fully correlated              \\
    \hline
    \multicolumn{10}{l}{\textbf{Normalisation, }$\PZ$ \textbf{production} }\\
    ZL,ZJ       & 4\% & 4\% & 4\% & 4\% & 4\%  & 4\% & 4\% & 4\% & Fully correlated              \\
    ZTT         &  & 4\% & 4\% & 4\% &  & 4\% & 4\% & 4\% & Fully correlated              \\
   \multicolumn{10}{l}{ $Z\to\Pgt\Pgt/Z\to\Pgm\Pgm$\textbf{ extrapolation} } \\
    ZTT         & 3\% & 3\% & 3\% & 3\% & 5\% & 5\% & 5\% & 5\% & Fully uncorrelated           \\
    \multicolumn{10}{l}{\textbf{Normalisation,} $\cPqt\cPaqt$}\\
    TTT,TTJ     & 6\% & 6\% & 6\% & 6\%   & 6\% & 6\% &6\% & 6\%  & Fully correlated                \\
    \multicolumn{10}{l}{\textbf{Normalisation, di-boson} } \\
    VVT,VVJ     & 5\% & 5\% & 5\% & 5\%   & 5\% & 5\% & 5\% & 5\% & Fully correlated         \\
   \multicolumn{10}{l}{ \textbf{Normalisation,} $\PZ\to\Pe\Pe$: $\Pe$ \textbf{misidentified as} $\tau_{h}$ }\\
    ZL     & 30\%  & 30\% & 30\% & 30\%   & 30\% & 30\% & 30\% &30\% & Corr. between cats, uncorr. between chns       \\
    \multicolumn{10}{l}{\textbf{Normalisation,} $\PZ$\textbf{+jets : jet misidentified as} $\tau_{h}$ } \\
    ZJ,TTJ     & 20\%  & 20\% &20\% &20\%      & 20\% & 20\% &20\% &20\%  & Corr. between cats, uncorr. between chns       \\
    \hline
    \multicolumn{10}{l}{\textbf{W OS/SS ratio (stat) } } \\
    W & 2\% & 2\% & & &14\% &14\% & & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W OS/SS ratio (syst) } }\\
    W & 8\% & 8\% & & &10\% &10\% & & &Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W low/high} $m_{T}$ \textbf{ratio (stat)}}\\
    W & 2\% & & 2 \% & &17\% & & 17\% & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W low/high} $m_{T}$ \textbf{ratio (syst)} }\\
    W & 20\% & &20\% & & 20\% & & 20\% & & Fully uncorrelated \\
    %\multicolumn{10}{l}{\textbf{W b-tag extrap (stat)}}\\
    %W &  & &  & &11\% &14\% & 21\% & 16\%& Fully uncorrelated \\
    %\multicolumn{10}{l}{\textbf{W b-tag extrap (syst)} }\\
    %W &  & & & & 10\% & & 10\% & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{QCD OS/SS ratio (syst) }}\\
    QCD & 12\% & 12\% & & & 60\% & 60\% & & &Fully uncorrelated \\
    \hline
\end{tabular}
}
\label{tab:SystematicUncertainties_et}
\end{center}
\end{table}

\begin{table}[!h]
\begin{center}
\caption{Systematic uncertainties that affect the estimated number of signal
or background events in the $\tau_{h} \tau_{h}$ channel.Some of the uncertainties
are different for different processes, where this is the case ranges are given.
Uncertainties marked "Fully correlated" are correlated between the different
decay channels and event categories, those marked "Fully uncorrelated" are
uncorrelated between the different decay channels and event categories.}
{\scriptsize
\begin{tabular}{l|cc|p{3cm}}
     & \multicolumn{2}{|c}{Event yield uncertainty by event category} & \\
    \hline
    Process & No b-tag & B-tag & Correlation                   \\
    \hline
    \multicolumn{4}{l}{\textbf{Integrated luminosity 13}\TeV}\\
    All from MC      & 6.2\%      & 6.2\%  & Fully correlated                           \\
    \hline
    \multicolumn{4}{l}{\textbf{Jet energy scale}} \\
    All from MC      & 0--4\%      & 0--9\%  & Fully correlated                    \\
    \hline
    \multicolumn{4}{l}{\MET \textbf{scale}} \\
    ZTT,W     & 2\%     & 2\% & Correlated between chn/cat,                          \\
    TTT,TTJ,VVT,VVJ     & 2\%     & 2\% & ZTT/W uncorrelated from TTT,TTJ,VVT,VVJ                          \\
    \hline
    \multicolumn{4}{l}{\MET \textbf{resolution}} \\
    ZTT,W     & 2\%     & 2\% & Correlated between chn/cat,                          \\
    TTT,TTJ,VVT,VVJ     & 2\%     & 2\% & ZTT/W uncorrelated from TTT,TTJ,VVT,VVJ                          \\
    \hline
    \multicolumn{4}{l}{Tau-lepton identification and trigger} \\
    ZTT,TTT,VVT, signal         & 10\%    & 10\%  & Fully correlated                      \\
    ZTT,TTT,VVT, signal         & 9.2\%     & 9.2\%   & Correlated between categories, uncorrelated between channels \\
    \hline
    \multicolumn{4}{l}{\textbf{B-tagging efficiency} }\\
    All from MC   & 0--4\%     & 0--11\%  & Fully correlated                  \\
    \hline
   \multicolumn{4}{l}{\textbf{B-mistag rate}} \\
    All from MC    &  0--1\%    & 0--12\%  & Fully correlated                  \\
    \hline
    \multicolumn{4}{l}{\textbf{Tau-lepton energy scale}} \\
    ZTT,TTT,VVT, signal       & shape      & shape & Correlated between categories, uncorrelated between channels                       \\
    \hline
    \multicolumn{4}{l}{\textbf{High} $P_{T}$ $\tau_h$ \textbf{ID efficiency}} \\
    ZTT,TTT,VVT, signal       & shape      & shape & Correlated between categories, uncorrelated between channels   \\
    \hline
    \multicolumn{4}{l}{\textbf{DY reweighting } } \\
    ZTT       & shape      & shape & Fully correlated                       \\
    \hline
    \multicolumn{4}{l}{\textbf{Top quark} $P_{T}$ \textbf{reweighting } } \\
    TTT,TTJ & shape & shape & Fully correlated                  \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, }$\PZ$ \textbf{production}}\\
    ZL,ZJ       & 4\%      & 4\%  & Fully correlated                    \\
    \hline
    \multicolumn{4}{l}{$Z\to\Pgt\Pgt/Z\to\Pgm\Pgm$\textbf{: extrapolation}}\\
    ZTT         & 3\%        & 5\%  & Fully uncorrelated                   \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, }$\cPqt\cPaqt$}\\
    TTT,TTJ        & 6\%       & 6\%  & Fully correlated                       \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, di-boson}}\\
    VVT,VVJ &  5\%        & 5\%       & Fully correlated                        \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, }$\PZ\to\Pe\Pe$: $\Pe$\textbf{ misidentified as }$\tau_{h}$}\\
    ZL & 10\%     & 10\%       & Correlated between categories, uncorrelated between channels                        \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, }$\PZ$\textbf{+jets : jet misidentified as} $\tau_{h}$}\\
    ZJ,W,TTJ & 20\%     & 20\%       & Correlated bewteen categories, uncorrelated between channels     \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, QCD multijet (stat) }} \\
    QCD  & 2\% & 20\%  & Fully uncorrelated\\
    \multicolumn{4}{l}{\textbf{Normalisation, QCD multijet (syst) }} \\
    QCD  & 12\% & 14\%  & Fully uncorrelated\\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, W + jets }}\\
    W & 4\% & 4\% & Fully correlated \\
\end{tabular}}
\label{tab:SystematicUncertainties_tt}
\end{center}
\end{table}

\begin{table}[!h]
\begin{center}
\caption{Systematic uncertainties that affect the estimated number of signal
or background events in the $e\mu$ channel.Some of the uncertainties
are different for different processes, where this is the case ranges are given.
Uncertainties marked "Fully correlated" are correlated between the different
decay channels and event categories, those marked "Fully uncorrelated" are
uncorrelated between the different decay channels and event categories.}
{\scriptsize
\begin{tabular}{l|cc|p{3cm}}
     & \multicolumn{2}{|c}{Event yield uncertainty by event category} &   \\
    \hline
    \textbf{Process}
    &  \textbf{No b-tag} & \textbf{B-tag} & \textbf{Correlation}                   \\
    \hline
    \multicolumn{4}{l}{\textbf{Integrated luminosity 13}\TeV}\\
    All from MC      & 6.2\%      & 6.2\% & Fully correlated                            \\
    \hline
    \multicolumn{4}{l}{\textbf{Jet energy scale}}\\
    All from MC   & 0--1\% & 1--69\% &Fully correlated \\
    \hline
    \multicolumn{4}{l}{\MET \textbf{scale}} \\
    ZTT,W    & 2\%     & 2\% & Correlated between chn/cat,                          \\
    TT,VV    & 2\%     & 2\% & ZTT,W uncorrelated from TT,VV \\
    \hline
   \multicolumn{4}{l}{ \MET \textbf{resolution}} \\
    ZTT,W & 2 \%    & 2\%  & Correlated between chn/cat,\\
    TT,VV & 2\%     & 2\%  & ZTT,W uncorrelated from TT,VV\\
    \hline
    \multicolumn{4}{l}{\textbf{Electron identification and trigger}}\\
    All from MC       & 2\%        & 2\% & Fully correlated                              \\
    \hline
    \multicolumn{4}{l}{\textbf{Muon identification and trigger}  }\\
    All from MC      & 2\%     & 2\%  & Fully correlated                      \\
    \hline
    \multicolumn{4}{l}{\textbf{B-tagging efficiency}} \\
    All from MC     & 0--3\%     & 1--3\%  & Fully correlated                  \\
    \hline
    \multicolumn{4}{l}{\textbf{B-mistag rate}}\\
    All from MC      &     & 0--10\% & Fully correlated                    \\
    \hline
    \multicolumn{4}{l}{\textbf{Electron energy scale}} \\
    ZTT,signal      & shape      & shape  & Fully correlated                \\
    \hline
    \multicolumn{4}{l}{\textbf{DY reweighting }}\\
     ZTT       & shape      & shape & Fully correlated                       \\
   \hline
    \multicolumn{4}{l}{\textbf{Top quark} $P_{T}$ \textbf{reweighting}}\\
    TT & shape & shape  & Fully correlated                 \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, }$\PZ$ \textbf{production} }\\
    ZLL      & 4\%      & 4\%  & Fully correlated                   \\
    \hline
    \multicolumn{4}{l}{$Z\to\Pgt\Pgt/Z\to\Pgm\Pgm$\textbf{: extrapolation} }\\
     ZTT        & 3\%        & 5\%   & Fully uncorrelated                      \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, }$\cPqt\cPaqt$}\\
    TT        & 6\%       & 6\% & Fully correlated                        \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, di-boson}} \\
    VV        & 5\%       & 5\% & Fully correlated                       \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, QCD multijet }}\\
    QCD & 23\% & 34\% & Fully uncorrelated\\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, W + jets }}\\
    W+jets & 4\% & 4\% & Fully correlated\\
\end{tabular}}
\label{tab:SystematicUncertainties_em}
\end{center}
\end{table}

\begin{table}[!h]
\begin{center}
\caption{Systematic uncertainties that affect the estimated number of
or background events in the $Z\rightarrow\mu\mu$ channel.Some of the uncertainties
are different for different processes, where this is the case ranges are given.
Uncertainties marked "Fully correlated" are correlated between the different
decay channels and event categories, those marked "Fully uncorrelated" are
uncorrelated between the different decay channels and event categories.}
{\scriptsize
\begin{tabular}{l|cc|p{3cm}}
     & \multicolumn{2}{|c}{Event yield uncertainty by event category} &   \\
    \hline
    \textbf{Process}
    &  \textbf{No b-tag} & \textbf{B-tag} & \textbf{Correlation}                   \\
    \hline
    \multicolumn{4}{l}{\textbf{Integrated luminosity 13}\TeV}\\
    All from MC (not ZL)     & 6.2\%      & 6.2\% & Fully correlated                            \\
    \hline
    \multicolumn{4}{l}{\textbf{Jet energy scale}}\\
    All from MC   & 0--1\% & 1--69\% &Fully correlated \\
    \hline
    \multicolumn{4}{l}{\MET \textbf{scale}} \\
    ZTT,W    & 2\%     & 2\% & Correlated between chn/cat,                          \\
    TT,VV    & 2\%     & 2\% & ZTT,W uncorrelated from TT,VV \\
    \hline
   \multicolumn{4}{l}{ \MET \textbf{resolution}} \\
    ZTT,W & 2 \%    & 2\%  & Correlated between chn/cat,\\
    TT,VV & 2\%     & 2\%  & ZTT,W uncorrelated from TT,VV\\
    \hline
    \multicolumn{4}{l}{\textbf{Muon identification and trigger}}\\
    All from MC       & 2--4\%        & 2--4\% & Fully correlated                              \\
    \hline
    \multicolumn{4}{l}{\textbf{B-tagging efficiency}} \\
    All from MC     & 0--3\%     & 1--3\%  & Fully correlated                  \\
    \hline
    \multicolumn{4}{l}{\textbf{B-mistag rate}}\\
    All from MC      &     & 0--10\% & Fully correlated                    \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, }$\cPqt\cPaqt$}\\
    TT        & 6\%       & 6\% & Fully correlated                        \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, di-boson}} \\
    VV        & 5\%       & 5\% & Fully correlated                       \\
    \hline
    \multicolumn{4}{l}{\textbf{Normalisation, W + jets }}\\
    W+jets & 4\% & 4\% & Fully correlated\\
\end{tabular}}
\label{tab:SystematicUncertainties_zmm}
\end{center}
\end{table}