### 2017-01-12 MSSM/UncOv.tex
Fixing tables in MSSM section

\begin{table}[!h]
\begin{center}
\caption{\footnotesize Systematic uncertainties that affect the estimated number of signal
or background events in the $\mu \tau_{h}$ channel.} 
 {\tiny
 \begin{tabular}{p{2cm}|p{1cm}p{1cm}p{1cm}p{1cm}|p{1cm}p{1cm}p{1cm}p{1cm}|p{3cm}}
\toprule
     & \multicolumn{8}{|c}{Event yield uncertainty by event category} &  \\
    \midrule
\textbf{ Process }
    & \multicolumn{4}{|c}{\textbf{No b-tag}} & \multicolumn{4}{|c}{\textbf{B-tag}} & \textbf{Correlation}           \\
     & SR & OS high $m_{T}$ & SS low $m_{T}$ & SS high $m_{T}$ & SR & OS high $m_{T}$ & SS low $m_{T}$ & SS high $m_{T}$ & \\
    \midrule
    \multicolumn{10}{l}{\textbf{Integrated luminosity 13}\TeV} \\
    All from MC    & 6.2\%  &6.2\%  & 6.2\%    & 6.2\% & 6.2\% & 6.2\% & 6.2\% & 6.2\% & Fully correlated\\
    \midrule
    \multicolumn{10}{l}{\textbf{Jet energy scale }}\\
    All from MC & 0--1\% & 0--2\% & 0--1\% & 0--1\%& 0--12\% & 3--11\% & 0--15\% & 0--14\% & Fully correlated \\
    \midrule
    \multicolumn{10}{l}{\MET \textbf{scale} }\\
    ZTT,W     & 2\% & & 2\% & &2\% &  & 2\% & & Corr. between chn/cat; TTT,TTJ, VVT,VVJ                         \\
    TTT,TTJ,VVT,VVJ & 2\% & & 2\% & &2\% & & 2\% & & uncorr. from ZTT,W \\
    \midrule
    \multicolumn{4}{l}{\MET \textbf{resolution}} \\
    ZTT,W     & 2\% & & 2\% & &2\% &  & 2\% & & Corr. between chn/cat; TTT,TTJ,VVT,VVJ                         \\
    TTT,TTJ,VVT,VVJ & 2\% & & 2\% & &2\% & & 2\% & & uncorr. from ZTT,W \\
    \midrule
    \multicolumn{10}{l}{\textbf{Muon identification and trigger} }\\
    All from MC & 2\% & 2\% &2\% &2\%  & 2\%  & 2\% & 2\% & 2\% & Fully correlated      \\
    \midrule
    \multicolumn{10}{l}{\textbf{Tau-lepton identification}}\\
    ZTT,TTT,VVT, signal     & 5\% & 5\% & 5\% & 5\%   & 5\%  & 5\% & 5\% & 5\% & Fully correlated \\
    ZTT,TTT,VVT, signal     & 3\% & 3\% & 3\% & 3\%   & 3\%  & 3\% & 3\% & 3\% & Cats:C,chns:U    \\
    \midrule
    \multicolumn{10}{l}{\textbf{B-tagging efficiency} }\\
    All from MC & 0--3\% & 0--3\% & 0--3\% & 0--2\% & 1--3\% &  0--3\%& 0--3\% & 0--3\%& Fully correlated\\
    W &  &  &  &  & 4\% & 4\% & 6\% & 6\%& Fully correlated\\
    \midrule
    \multicolumn{10}{l}{\textbf{B-mistag rate } }\\
    All from MC & & & & & 1--7\% & & & & Fully correlated\\
    \midrule
    \multicolumn{10}{l}{\textbf{Tau-lepton energy scale}}\\
    ZTT,TTT,VVT, signal     & shape & shape & shape & shape  & shape & shape & shape & shape & Cats:C,chns:U   \\
    \midrule
    \multicolumn{10}{l}{\textbf{High} $P_{T}$ $\tau_h$\textbf{ ID efficiency } } \\
    ZTT,TTT,VVT, signal    & shape & shape & shape & shape  & shape & shape & shape & shape & Cats:C,chns:U   \\
    \midrule
    \multicolumn{10}{l}{\textbf{Top quark} $P_{T}$ \textbf{reweighting} }\\
    TTT,TTJ  & shape & shape & shape & shape & shape & shape & shape & shape & Fully correlated    \\
    \midrule
    \multicolumn{10}{l}{\textbf{DY reweighting } }\\
    ZTT       & shape & shape & shape & shape  & shape & shape & shape & shape &Fully correlated              \\
    \midrule
    \multicolumn{10}{l}{\textbf{Jet}$\rightarrow\tau_h$ \textbf{fake rate reweighting } }\\
    W         & shape &  &  &   & shape &  &  &  &Fully correlated              \\
    \midrule
    \multicolumn{10}{l}{\textbf{Normalisation, }$\PZ$ \textbf{production} }\\
    ZL,ZJ       & 4\% & 4\% & 4\% & 4\% & 4\%  & 4\% & 4\% & 4\% & Fully correlated              \\
    ZTT         &  & 4\% & 4\% & 4\% &  & 4\% & 4\% & 4\% & Fully correlated              \\
   \multicolumn{10}{l}{ $Z\to\Pgt\Pgt$\textbf{: $Z\to\Pgt\Pgt / Z\to\Pgm\Pgm$ extrapolation } } \\
    ZTT         & 3\% & 3\% & 3\% & 3\% & 5\% & 5\% & 5\% & 5\% & Fully uncorrelated           \\
    \multicolumn{10}{l}{\textbf{Normalisation,} \ttbar}\\
    TTT,TTJ     & 6\% & 6\% & 6\% & 6\%   & 6\% & 6\% &6\% & 6\%  & Fully correlated                \\
    \multicolumn{10}{l}{\textbf{Normalisation, di-boson} } \\
    VVT,VVJ     & 5\% & 5\% & 5\% & 5\%   & 5\% & 5\% & 5\% & 5\% & Fully correlated         \\
   \multicolumn{10}{l}{ \textbf{Normalisation,} $\PZ\to\Pgm\Pgm$: $\Pgm$ \textbf{misidentified as} $\tau_{h}$ }\\
    ZL     & 30\%  & 30\% & 30\% & 30\%   & 30\% & 30\% & 30\% &30\% &Cats:C,chns:U        \\
    \multicolumn{10}{l}{\textbf{Normalisation,} $\PZ$\textbf{+jets : jet misidentified as} $\tau_{h}$ } \\
    ZJ,TTJ     & 20\%  & 20\% &20\% &20\%      & 20\% & 20\% &20\% &20\%  & Cats:C,chns:U  \\
    \midrule
    \multicolumn{10}{l}{\textbf{W OS/SS ratio (stat) } } \\
    W & 2\% & 2\% & & &11\% &11\% & & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W OS/SS ratio (syst) } }\\
    W & 8\% & 8\% & & &10\% &10\% & & &Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W low/high} $m_{T}$ \textbf{ratio (stat)}}\\
    W & 2\% & & 2 \% & &14\% & & 14\% & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W low/high} $m_{T}$ \textbf{ratio (syst)} }\\
    W & 20\% & &20\% & & 20\% & & 20\% & & Fully uncorrelated \\
    %\multicolumn{10}{l}{\textbf{W b-tag extrap (stat)}}\\
    %W &  & &  & &12\% & 8\%& 22\% & 14\%& Fully uncorrelated \\
    %\multicolumn{10}{l}{\textbf{W b-tag extrap (syst)} }\\
    %W &  & & & & 10\% & & 10\% & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{QCD OS/SS ratio (syst) }}\\
    QCD & 4\% & 4\% & & & 60\% & 60\% & & &Fully uncorrelated \\
\bottomrule
\end{tabular}
}
\label{tab:SystematicUncertainties_mt}
\end{center}
\end{table}

\begin{table}[!h]
\begin{center}
\caption{\footnotesize Systematic uncertainties that affect the estimated number of signal
or background events in the \etau channel.}
 {\tiny
 \begin{tabular}{p{2cm}|p{1cm}p{1cm}p{1cm}p{1cm}|p{1cm}p{1cm}p{1cm}p{1cm}|p{3cm}}
\toprule
     & \multicolumn{8}{|c}{Event yield uncertainty by event category} &  \\
    \midrule
\textbf{ Process }
    & \multicolumn{4}{|c}{\textbf{No b-tag}} & \multicolumn{4}{|c}{\textbf{B-tag}} & \textbf{Correlation}           \\
     & SR & OS high $m_{T}$ & SS low $m_{T}$ & SS high $m_{T}$ & SR & OS high $m_{T}$ & SS low $m_{T}$ & SS high $m_{T}$ & \\
    \midrule
    \multicolumn{10}{l}{\textbf{Integrated luminosity 13}\TeV} \\
    All from MC    & 6.2\%  & 6.2\%  & 6.2\%    & 6.2\% & 6.2\% & 6.2\% & 6.2\% & 6.2\% & Fully correlated\\
    \midrule
    \multicolumn{10}{l}{\textbf{Jet energy scale }}\\
    All from MC & 0--1\% & 0--1\% & 0--2\% & 0--1\%& 1--14\% & 0--12\% & 0--15\% & 0--15\% & Fully correlated \\
    \midrule
    \multicolumn{10}{l}{\MET \textbf{scale}}\\
    ZTT,W     & 2\% & & 2\% & &2\% &  & 2\% & & Corr. between chn/cat, TTT,TTJ,VVT,VVJ                         \\
    TTT,TTJ,VVT,VVJ & 2\% & & 2\% & &2\% & & 2\% & & uncorr. from ZTT,W \\
    \midrule
    \multicolumn{10}{l}{\MET \textbf{resolution}} \\
    ZTT,W     & 2\% & & 2\% & &2\% &  & 2\% & & Corr. between chn/cat, TTT,TTJ,VVT,VVJ          \\
    TTT,TTJ,VVT,VVJ & 2\% & & 2\% & &2\% & & 2\% & & uncorr. from ZTT,W \\
    \midrule
    \multicolumn{10}{l}{\textbf{Electron identification and trigger} }\\
    All from MC & 2\% & 2\% &2\% &2\%  & 2\%  & 2\% & 2\% & 2\% & Fully correlated      \\
    \midrule
    \multicolumn{10}{l}{\textbf{Tau-lepton identification}}\\
    ZTT,TTT,VVT, signal     & 5\% & 5\% & 5\% & 5\%   & 5\%  & 5\% & 5\% & 5\% & Fully correlated \\
    ZTT,TTT,VVT, signal     & 3\% & 3\% & 3\% & 3\%   & 3\%  & 3\% & 3\% & 3\% & Cats:C,chns:U    \\
    \midrule
    \multicolumn{10}{l}{\textbf{B-tagging efficiency} }\\
    All from MC & 0--3\% & 0--3\% & 0--4\% & 0--4\% & 1--4\% & 0--2\% & 0--2\% & 0--2\% & Fully correlated\\
    W &  &  &  &  & 5\% & 2\% & 10\% & 3\% & Fully correlated\\
    \midrule
    \multicolumn{10}{l}{\textbf{B-mistag rate } }\\
    All from MC & & & & & 0--3\% & & & & Fully correlated\\
    \midrule
    \multicolumn{10}{l}{\textbf{Tau-lepton energy scale}}\\
    ZTT,TTT,VVT, signal     & shape & shape & shape & shape  & shape & shape & shape & shape & Cats:C,chns:U   \\
    \midrule
    \multicolumn{10}{l}{\textbf{High} $P_{T}$ $\tau_h$\textbf{ ID efficiency } } \\
    ZTT,TTT,VVT, signal    & shape & shape & shape & shape  & shape & shape & shape & shape & Cats:C,chns:U   \\
    \midrule
    \multicolumn{10}{l}{\textbf{Top quark} $P_{T}$ \textbf{reweighting} }\\
    TTT,TTJ  & shape & shape & shape & shape & shape & shape & shape & shape & Fully correlated    \\
    \midrule
    \multicolumn{10}{l}{\textbf{DY reweighting } }\\
    ZTT       & shape & shape & shape & shape  & shape & shape & shape & shape &Fully correlated              \\
    \midrule
    \multicolumn{10}{l}{\textbf{Jet}$\rightarrow\tau_h$ \textbf{fake rate reweighting } }\\
    W         & shape &  &  &   & shape &  &  &  &Fully correlated              \\
    \midrule
    \multicolumn{10}{l}{\textbf{Normalisation, }$\PZ$ \textbf{production} }\\
    ZL,ZJ       & 4\% & 4\% & 4\% & 4\% & 4\%  & 4\% & 4\% & 4\% & Fully correlated              \\
    ZTT         &  & 4\% & 4\% & 4\% &  & 4\% & 4\% & 4\% & Fully correlated              \\
   \multicolumn{10}{l}{ $Z\to\Pgt\Pgt/Z\to\Pgm\Pgm$\textbf{ extrapolation} } \\
    ZTT         & 3\% & 3\% & 3\% & 3\% & 5\% & 5\% & 5\% & 5\% & Fully uncorrelated           \\
    \multicolumn{10}{l}{\textbf{Normalisation,} \ttbar}\\
    TTT,TTJ     & 6\% & 6\% & 6\% & 6\%   & 6\% & 6\% &6\% & 6\%  & Fully correlated                \\
    \multicolumn{10}{l}{\textbf{Normalisation, di-boson} } \\
    VVT,VVJ     & 5\% & 5\% & 5\% & 5\%   & 5\% & 5\% & 5\% & 5\% & Fully correlated         \\
   \multicolumn{10}{l}{ \textbf{Normalisation,} $\PZ\to\Pe\Pe$: $\Pe$ \textbf{misidentified as} $\tau_{h}$ }\\
    ZL     & 30\%  & 30\% & 30\% & 30\%   & 30\% & 30\% & 30\% &30\% & Cats:C,chns:U       \\
    \multicolumn{10}{l}{\textbf{Normalisation,} $\PZ$\textbf{+jets : jet misidentified as} $\tau_{h}$ } \\
    ZJ,TTJ     & 20\%  & 20\% &20\% &20\%      & 20\% & 20\% &20\% &20\%  & Cats:C,chns:U       \\
    \midrule
    \multicolumn{10}{l}{\textbf{W OS/SS ratio (stat) } } \\
    W & 2\% & 2\% & & &14\% &14\% & & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W OS/SS ratio (syst) } }\\
    W & 8\% & 8\% & & &10\% &10\% & & &Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W low/high} $m_{T}$ \textbf{ratio (stat)}}\\
    W & 2\% & & 2 \% & &17\% & & 17\% & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{W low/high} $m_{T}$ \textbf{ratio (syst)} }\\
    W & 20\% & &20\% & & 20\% & & 20\% & & Fully uncorrelated \\
    %\multicolumn{10}{l}{\textbf{W b-tag extrap (stat)}}\\
    %W &  & &  & &11\% &14\% & 21\% & 16\%& Fully uncorrelated \\
    %\multicolumn{10}{l}{\textbf{W b-tag extrap (syst)} }\\
    %W &  & & & & 10\% & & 10\% & & Fully uncorrelated \\
    \multicolumn{10}{l}{\textbf{QCD OS/SS ratio (syst) }}\\
    QCD & 12\% & 12\% & & & 60\% & 60\% & & &Fully uncorrelated \\
    \bottomrule
\end{tabular}
}
\label{tab:SystematicUncertainties_et}
\end{center}
\end{table}

\begin{table}[!h]
\begin{center}
\caption{Systematic uncertainties that affect the estimated number of signal
or background events in the \tautau channel.}
{\scriptsize
\begin{tabular}{l|cc|p{5cm}}
\toprule
     & \multicolumn{2}{|c}{Event yield uncertainty by event category} & \\
    \midrule
    Process & No b-tag & B-tag & Correlation                   \\
    \midrule
    \multicolumn{4}{l}{\textbf{Integrated luminosity 13}\TeV}\\
    All from MC      & 6.2\%      & 6.2\%  & Fully correlated                           \\
    \midrule
    \multicolumn{4}{l}{\textbf{Jet energy scale}} \\
    All from MC      & 0--4\%      & 0--9\%  & Fully correlated                    \\
    \midrule
    \multicolumn{4}{l}{\MET \textbf{scale}} \\
    ZTT,W     & 2\%     & 2\% & Correlated between chn/cat,                          \\
    TTT,TTJ,VVT,VVJ     & 2\%     & 2\% & ZTT/W uncorrelated from TTT,TTJ,VVT,VVJ                          \\
    \midrule
    \multicolumn{4}{l}{\MET \textbf{resolution}} \\
    ZTT,W     & 2\%     & 2\% & Correlated between chn/cat,                          \\
    TTT,TTJ,VVT,VVJ     & 2\%     & 2\% & ZTT/W uncorrelated from TTT,TTJ,VVT,VVJ                          \\
    \midrule
    \multicolumn{4}{l}{Tau-lepton identification and trigger} \\
    ZTT,TTT,VVT, signal         & 10\%    & 10\%  & Fully correlated                      \\
    ZTT,TTT,VVT, signal         & 9.2\%     & 9.2\%   & Correlated between categories, uncorrelated between channels \\
    \midrule
    \multicolumn{4}{l}{\textbf{B-tagging efficiency} }\\
    All from MC   & 0--4\%     & 0--11\%  & Fully correlated                  \\
    \midrule
   \multicolumn{4}{l}{\textbf{B-mistag rate}} \\
    All from MC    &  0--1\%    & 0--12\%  & Fully correlated                  \\
    \midrule
    \multicolumn{4}{l}{\textbf{Tau-lepton energy scale}} \\
    ZTT,TTT,VVT, signal       & shape      & shape & Correlated between categories, uncorrelated between channels                       \\
    \midrule
    \multicolumn{4}{l}{\textbf{High} $P_{T}$ $\tau_h$ \textbf{ID efficiency}} \\
    ZTT,TTT,VVT, signal       & shape      & shape & Correlated between categories, uncorrelated between channels   \\
    \midrule
    \multicolumn{4}{l}{\textbf{DY reweighting } } \\
    ZTT       & shape      & shape & Fully correlated                       \\
    \midrule
    \multicolumn{4}{l}{\textbf{Top quark} $P_{T}$ \textbf{reweighting } } \\
    TTT,TTJ & shape & shape & Fully correlated                  \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, }$\PZ$ \textbf{production}}\\
    ZL,ZJ       & 4\%      & 4\%  & Fully correlated                    \\
    \midrule
    \multicolumn{4}{l}{$Z\to\Pgt\Pgt/Z\to\Pgm\Pgm$\textbf{: extrapolation}}\\
    ZTT         & 3\%        & 5\%  & Fully uncorrelated                   \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, }\ttbar}\\
    TTT,TTJ        & 6\%       & 6\%  & Fully correlated                       \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, di-boson}}\\
    VVT,VVJ &  5\%        & 5\%       & Fully correlated                        \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, }$\PZ\to\Pe\Pe$: $\Pe$\textbf{ misidentified as }$\tau_{h}$}\\
    ZL & 10\%     & 10\%       & Correlated between categories, uncorrelated between channels                        \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, }$\PZ$\textbf{+jets : jet misidentified as} $\tau_{h}$}\\
    ZJ,W,TTJ & 20\%     & 20\%       & Correlated bewteen categories, uncorrelated between channels     \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, QCD multijet (stat) }} \\
    QCD  & 2\% & 20\%  & Fully uncorrelated\\
    \multicolumn{4}{l}{\textbf{Normalisation, QCD multijet (syst) }} \\
    QCD  & 12\% & 14\%  & Fully uncorrelated\\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, W + jets }}\\
    W & 4\% & 4\% & Fully correlated \\
\bottomrule
\end{tabular}}
\label{tab:SystematicUncertainties_tt}
\end{center}
\end{table}

\begin{table}[!h]
\begin{center}
\caption{Systematic uncertainties that affect the estimated number of signal
or background events in the \emu channel.}
{\scriptsize
\begin{tabular}{l|cc|p{5cm}}
   \toprule
     & \multicolumn{2}{|c}{Event yield uncertainty by event category} &   \\
    \midrule
    \textbf{Process}
    &  \textbf{No b-tag} & \textbf{B-tag} & \textbf{Correlation}                   \\
    \midrule
    \multicolumn{4}{l}{\textbf{Integrated luminosity 13}\TeV}\\
    All from MC      & 6.2\%      & 6.2\% & Fully correlated                            \\
    \midrule
    \multicolumn{4}{l}{\textbf{Jet energy scale}}\\
    All from MC   & 0--1\% & 1--69\% &Fully correlated \\
    \midrule
    \multicolumn{4}{l}{\MET \textbf{scale}} \\
    ZTT,W    & 2\%     & 2\% & Correlated between chn/cat,                          \\
    TT,VV    & 2\%     & 2\% & ZTT,W uncorrelated from TT,VV \\
    \midrule
   \multicolumn{4}{l}{ \MET \textbf{resolution}} \\
    ZTT,W & 2 \%    & 2\%  & Correlated between chn/cat,\\
    TT,VV & 2\%     & 2\%  & ZTT,W uncorrelated from TT,VV\\
    \midrule
    \multicolumn{4}{l}{\textbf{Electron identification and trigger}}\\
    All from MC       & 2\%        & 2\% & Fully correlated                              \\
    \midrule
    \multicolumn{4}{l}{\textbf{Muon identification and trigger}  }\\
    All from MC      & 2\%     & 2\%  & Fully correlated                      \\
    \midrule
    \multicolumn{4}{l}{\textbf{B-tagging efficiency}} \\
    All from MC     & 0--3\%     & 1--3\%  & Fully correlated                  \\
    \midrule
    \multicolumn{4}{l}{\textbf{B-mistag rate}}\\
    All from MC      &     & 0--10\% & Fully correlated                    \\
    \midrule
    \multicolumn{4}{l}{\textbf{Electron energy scale}} \\
    ZTT,signal      & shape      & shape  & Fully correlated                \\
    \midrule
    \multicolumn{4}{l}{\textbf{DY reweighting }}\\
     ZTT       & shape      & shape & Fully correlated                       \\
   \midrule
    \multicolumn{4}{l}{\textbf{Top quark} $P_{T}$ \textbf{reweighting}}\\
    TT & shape & shape  & Fully correlated                 \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, }$\PZ$ \textbf{production} }\\
    ZLL      & 4\%      & 4\%  & Fully correlated                   \\
    \midrule
    \multicolumn{4}{l}{$Z\to\Pgt\Pgt/Z\to\Pgm\Pgm$\textbf{: extrapolation} }\\
     ZTT        & 3\%        & 5\%   & Fully uncorrelated                      \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, }\ttbar}\\
    TT        & 6\%       & 6\% & Fully correlated                        \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, di-boson}} \\
    VV        & 5\%       & 5\% & Fully correlated                       \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, QCD multijet }}\\
    QCD & 23\% & 34\% & Fully uncorrelated\\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, W + jets }}\\
    W+jets & 4\% & 4\% & Fully correlated\\
\bottomrule
\end{tabular}}
\label{tab:SystematicUncertainties_em}
\end{center}
\end{table}

\begin{table}[!h]
\begin{center}
\caption{Systematic uncertainties that affect the estimated number of
or background events in the $Z\rightarrow\mu\mu$ channel.}
{\scriptsize
\begin{tabular}{l|cc|p{5cm}}
   \toprule
     & \multicolumn{2}{|c}{Event yield uncertainty by event category} &   \\
    \midrule
    \textbf{Process}
    &  \textbf{No b-tag} & \textbf{B-tag} & \textbf{Correlation}                   \\
    \midrule
    \multicolumn{4}{l}{\textbf{Integrated luminosity 13}\TeV}\\
    All from MC (not ZL)     & 6.2\%      & 6.2\% & Fully correlated                            \\
    \midrule
    \multicolumn{4}{l}{\textbf{Jet energy scale}}\\
    All from MC   & 0--1\% & 1--69\% &Fully correlated \\
    \midrule
    \multicolumn{4}{l}{\MET \textbf{scale}} \\
    ZTT,W    & 2\%     & 2\% & Correlated between chn/cat,                          \\
    TT,VV    & 2\%     & 2\% & ZTT,W uncorrelated from TT,VV \\
    \midrule
   \multicolumn{4}{l}{ \MET \textbf{resolution}} \\
    ZTT,W & 2 \%    & 2\%  & Correlated between chn/cat,\\
    TT,VV & 2\%     & 2\%  & ZTT,W uncorrelated from TT,VV\\
    \midrule
    \multicolumn{4}{l}{\textbf{Muon identification and trigger}}\\
    All from MC       & 2--4\%        & 2--4\% & Fully correlated                              \\
    \midrule
    \multicolumn{4}{l}{\textbf{B-tagging efficiency}} \\
    All from MC     & 0--3\%     & 1--3\%  & Fully correlated                  \\
    \midrule
    \multicolumn{4}{l}{\textbf{B-mistag rate}}\\
    All from MC      &     & 0--10\% & Fully correlated                    \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, }\ttbar}\\
    TT        & 6\%       & 6\% & Fully correlated                        \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, di-boson}} \\
    VV        & 5\%       & 5\% & Fully correlated                       \\
    \midrule
    \multicolumn{4}{l}{\textbf{Normalisation, W + jets }}\\
    W+jets & 4\% & 4\% & Fully correlated\\
    \bottomrule
\end{tabular}}
\label{tab:SystematicUncertainties_zmm}
\end{center}
\end{table}