### 2017-02-12 frontmatter.tex
Committing some updates before edit round 1

%% Title
\titlepage[Imperial College London\\ Department of Physics]{
  A dissertation submitted to Imperial College London\\ for the degree of Doctor of Philosophy}

%\begin{abstract}[]
%The copyright of this thesis rests with the author and is made available under a
%Creative Commons Attribution Non-Commercial No Derivatives licence. Researchers
%are free to copy, distribute or transmit the thesis on the condition that they
%attribute it, that they do not use it for commercial purposes and that they do
%not alter, transform or build upon it. For any reuse or redistribution,
%researchers must make clear to others the licence terms of this work.
%\end{abstract}
%% Abstract
%\begin{abstract}%[\smaller \thetitle\\ \vspace*{1cm} \smaller {\theauthor}]
%  %\thispagestyle{empty}
%\end{abstract}

%% Declaration
%\begin{declaration}
%  \vspace*{1cm}
%  \begin{flushright}
%  \end{flushright}
%\end{declaration}

%% Acknowledgements
%\begin{acknowledgements}
%\end{acknowledgements}

%% Preface
%\begin{preface}
%\end{preface}
%% ToC
\tableofcontents
%\listoffigures
%\listoftables

%% Strictly optional!
%% I don't want a page number on the following blank page either.
\thispagestyle{empty}