### 2017-02-17 frontmatter.tex

%% Title
\titlepage[Imperial College London\\ Department of Physics]{
  A dissertation submitted to Imperial College London\\ for the degree of Doctor of Philosophy}

\begin{abstract}[]
The copyright of this thesis rests with the author and is made available under a
Creative Commons Attribution Non-Commercial No Derivatives licence. Researchers
are free to copy, distribute or transmit the thesis on the condition that they
attribute it, that they do not use it for commercial purposes and that they do
not alter, transform or build upon it. For any reuse or redistribution,
researchers must make clear to others the licence terms of this work.
\end{abstract}

% Abstract
\begin{abstract}%[\smaller \thetitle\\ \vspace*{1cm} \smaller {\theauthor}]
  %\thispagestyle{empty}
The \ac{CMS} detector is a general-purpose detector located at the Large Hadron
Collider at CERN. It was designed to test the predictions of the standard model 
and to search for the Higgs boson in addition to searches for new physics beyond
the standard model.
The analyses presented in this thesis are searches for neutral
Higgs bosons with tau leptons in the final state, and use proton-proton collision data recorded
by \acs{CMS} during 2012, 2015 and 2016. The searches are performed in the
context of the \ac{MSSM}. One of the analyses, using $19.7\,\invfb$ collected
during 2012 at $\sqrt{s}=8\,\TeV$, is a search for a heavy neutral Higgs boson
decaying into two $125\,\GeV$ \ac{MSSM} Higgs bosons with \ac{SM}-like properties,
with two b-quarks and two tau leptons in the final state. The other analyses use 
$2.3\,\invfb$ collected during 2015 and $12.9\,\invfb$ collected during 2016, both
at $\sqrt{s}=13\,\TeV$. These are searches for neutral \ac{MSSM}
Higgs bosons in the mass range of $90\,\GeV$--$3.2\,\TeV$ directly decaying
into pairs of tau leptons. 
No significant excess is observed in any of the searches. Upper limits at 95\% confidence 
level are set on the cross section times branching ratio for the Higgs production process
in consideration. In addition, limits are set in the \mA-\tanb~parameter space of 
several \ac{MSSM} benchmark scenarios, and in the \cosba-\tanb~parameter space of an example
of a two Higgs doublet model.

\end{abstract}

% Declaration
\begin{declaration}
  \vspace*{1cm}
  I declare that the work in this thesis is mine. Figures and results taken from other sources
include the appropriate reference in the text or figure caption. Figures labelled `CMS' are sourced
directly from CMS publications. The `CMS Preliminary' label indicates figures which
have been made public via a preliminary public document or a public website, but are not
included in a publication in a peer-reviewed journal. Figures labelled `CMS' or `CMS preliminary', 
including those made by myself, include the relevant reference in the figure caption.
The work for the analyses presented in chapters \ref{chap:hhh} and \ref{chap:mssm} was 
carried out as part of the CMS \Htotautau working group, and in collaboration with other members of the group.
For the results presented in chapter \ref{chap:hhh} I 
was responsible for the statistical inference. This included
the interpretation of the results in MSSM and two Higgs doublet model benchmark scenarios. The 
work is included in Ref. \cite{CMS-HIG-14-034}. For the results presented in chapter \ref{chap:mssm}
I was responsible for all stages of the analysis, including the optimisation
of object selection and event categorisation, as well as studies of background
methods for three out of four channels presented in the final version of the analysis,
evaluation of systematic uncertainties in all four channels, and production of the statistical
results, including model interpretations. This work is included in the most recent
preliminary result \cite{CMS-PAS-HIG-16-037}, as well as a predating preliminary
result using a smaller dataset \cite{CMS-PAS-HIG-16-006}.
The work presented in chapter \ref{sec:mssm_combination} has
not been made public, and draws on the results of \cite{CMS-PAS-HIG-16-037} and \cite{CMS-PAS-HIG-16-006}.

  \begin{flushright}
  Adinda de Wit
  \end{flushright}
\end{declaration}

%% Acknowledgements
%\begin{acknowledgements}
%\end{acknowledgements}

%% Preface
%\begin{preface}
%\end{preface}
%% ToC
\tableofcontents
%\listoffigures
%\listoftables

%% Strictly optional!
%% I don't want a page number on the following blank page either.
\thispagestyle{empty}