### 2017-02-24 backmatter.tex
More editing, plus additional cosmetic fixes from quick skim

%% You're recommended to use the eprint-aware biblio styles which
%% can be obtained from e.g. www.arxiv.org. The file mythesis.bib
%% is derived from the source using the SPIRES Bibtex service.
\bibliographystyle{lucas_unsrt}
\bibliography{references}

%% I prefer to put these tables here rather than making the
%% front matter seemingly interminable. No-one cares, anyway!
%\listoffigures
%\listoftables

\chapter*{List of Acronyms}
\addcontentsline{toc}{chapter}{List of Acronyms}
\begin{acronym}[CERN]
\acro{EM}[EM]{electromagnetic}
\acro{QFT}[QFT]{quantum field theory}
\acro{QCD}[QCD]{quantum chromodynamics}
\acro{QED}[QED]{quantum electrodynamics}
\acro{LEP}[LEP]{Large Electron-Positron}
\acro{MC}[MC]{Monte Carlo}
\acro{CKM}[CKM]{Cabibbo-Kobayashi-Maskawa}
\acro{CMS}[CMS]{Compact Muon Solenoid}
\acro{LHC}[LHC]{Large Hadron Collider}
\acro{CERN}[CERN]{the European Organisation for Nuclear Research }
\acro{PS}[PS]{Proton Synchrotron}
\acro{SPS}[SPS]{Super Proton Synchrotron}
\acro{PSB}[PSB]{Proton Synchrotron Booster}
\acro{LS1}[LS1]{Long Shutdown 1}
\acro{MSSM}[MSSM]{minimal supersymmetric standard model}
\acro{SM}[SM]{standard model}
\acro{SUSY}[SUSY]{supersymmetry}
\acro{2HDM}[2HDM]{two Higgs doublet model}
\acro{ECAL}[ECAL]{electromagnetic calorimeter}
\acro{HCAL}[HCAL]{hadron calorimeter}
\acro{TIB}[TIB]{Tracker Inner Barrel}
\acro{TID}[TID]{Tracker Inner Disk}
\acro{TOB}[TOB]{Tracker Outer Barrel}
\acro{TEC}[TEC]{Tracker EndCap}
\acro{EB}[EB]{ECAL Barrel}
\acro{EE}[EE]{ECAL Endcap}
\acro{HB}[HB]{Hadron Barrel Calorimeter}
\acro{HO}[HO]{Hadron Outer Calorimeter}
\acro{HE}[HE]{Hadron Endcap Calorimeter}
\acro{HF}[HF]{Hadron Forward Calorimeter}
\acro{BSM}[BSM]{beyond the standard model}
\acro{L1}[L1]{Level-1}
\acro{HLT}[HLT]{High-level trigger}
\acro{CSC}[CSC]{Cathode Strip Chamber}
\acro{DT}[DT]{Drift Tube chamber}
\acro{RPC}[RPC]{Resistive Plate Chamber}
\acro{DAQ}[DAQ]{Data Acquisition}
\acro{WLCG}[WLCG]{Worldwide LHC Computing Grid}
\acro{RF}[RF]{radio frequency}
\acro{PF}[PF]{Particle Flow}
\acro{BDT}[BDT]{boosted decision tree}
\acro{KF}[KF]{Kalman Filter}
\acro{CTF}[CTF]{Combinatorial Track Finder}
\acro{GSF}[GSF]{Gaussian Sum Filter}
%\acro{KF}[KF]{Kalman Filter}
\acro{DA}[DA]{Deterministic Annealing}
\acro{CHS}[CHS]{charged hadron subtraction}
\acro{CSV}[CSV]{Combined Secondary Vertex}
\acro{CL}[CL]{confidence level}
\acro{IVF}[IVF]{Inclusive Vertex Finder}
\acro{UE}[UE]{underlying event}
\acro{MPF}[MPF]{missing transverse energy projection fraction}
\end{acronym}

\appendix
\renewcommand{\chaptername}{Appendix}
\chapter{\texorpdfstring{Overview of systematic uncertainties in the MSSM \AHtotautau analysis}{Overview of systematic uncertainties in the MSSM A/H to tautau analysis}}
\label{appendix:uncerts}
An overview of the systematic uncertainties that are applied in the MSSM \AHtotautau analysis (chapter \ref{chap:mssm})
are summarised in this appendix. The uncertainties are given in
table \ref{tab:SystematicUncertainties_mt} for the $\mu\tau_h$
channel, in table \ref{tab:SystematicUncertainties_et} for the $e\tau_h$ channel,
in table \ref{tab:SystematicUncertainties_tt} for the $\tau_h\tau_h$ channel and
in table \ref{tab:SystematicUncertainties_em} for the $e\mu$ channel. Additionally,
the systematic uncertainties for the $Z\rightarrow\mumu$ control region are shown
in table \ref{tab:SystematicUncertainties_zmm}. Correlations are also indicated.
Some of the uncertainties
are different for different processes, where this is the case ranges are given.
Uncertainties marked "Fully correlated" are correlated between the different
decay channels and event categories, those marked "Fully uncorrelated" are
uncorrelated between the different decay channels and event categories. "Cats:C,chns:U"
indicates that the uncertainty is correlated between categories but uncorrelated between
the different decay channels.
\input{MSSM/UncOv}

%% If you have time and interest to generate a (decent) index,
%% then you've clearly spent more time on the write-up than the 
%% research ;-)
%\printindex