### 2017-03-03 backmatter.tex
Own final read through

%% You're recommended to use the eprint-aware biblio styles which
%% can be obtained from e.g. www.arxiv.org. The file mythesis.bib
%% is derived from the source using the SPIRES Bibtex service.
\bibliographystyle{lucas_unsrt}
\bibliography{references}

%% I prefer to put these tables here rather than making the
%% front matter seemingly interminable. No-one cares, anyway!
%\listoffigures
%\listoftables

\chapter*{List of acronyms}
\addcontentsline{toc}{chapter}{List of acronyms}
\markboth{LIST OF ACRONYMS}{}
\begin{acronym}
\acro{2HDM}[2HDM]{two Higgs doublet model}
\acro{BDT}[BDT]{boosted decision tree}
\acro{BSM}[BSM]{beyond the standard model}
\acro{CERN}[CERN]{the European Organisation for Nuclear Research }
\acro{CHS}[CHS]{charged hadron subtraction}
\acro{CKM}[CKM]{Cabibbo-Kobayashi-Maskawa}
\acro{CL}[CL]{confidence level}
\acro{CMS}[CMS]{Compact Muon Solenoid}
\acro{CSC}[CSC]{cathode strip chamber}
\acro{CSV}[CSV]{combined secondary certex}
\acro{CTF}[CTF]{combinatorial track finder}
\acro{DA}[DA]{deterministic annealing}
\acro{DAQ}[DAQ]{aata acquisition}
\acro{DT}[DT]{drift tube chamber}
\acro{EB}[EB]{ECAL barrel}
\acro{ECAL}[ECAL]{electromagnetic calorimeter}
\acro{EE}[EE]{ECAL endcap}
\acro{EM}[EM]{electromagnetic}
\acro{GSF}[GSF]{gaussian sum filter}
\acro{HB}[HB]{hadron barrel calorimeter}
\acro{HCAL}[HCAL]{hadron calorimeter}
\acro{HE}[HE]{hadron endcap calorimeter}
\acro{HF}[HF]{hadron forward calorimeter}
\acro{HLT}[HLT]{High-level trigger}
\acro{HO}[HO]{hadron outer calorimeter}
\acro{IVF}[IVF]{inclusive vertex finder}
\acro{KF}[KF]{Kalman filter}
\acro{L1}[L1]{Level-1}
\acro{LEP}[LEP]{Large Electron-Positron}
\acro{LHC}[LHC]{Large Hadron Collider}
\acro{LS1}[LS1]{Long Shutdown 1}
\acro{MC}[MC]{Monte Carlo}
\acro{MPF}[MPF]{missing transverse energy projection fraction}
\acro{MSSM}[MSSM]{minimal supersymmetric standard model}
\acro{QCD}[QCD]{quantum chromodynamics}
\acro{PF}[PF]{particle flow}
\acro{PS}[PS]{Proton Synchrotron}
\acro{PSB}[PSB]{Proton Synchrotron Booster}
\acro{QED}[QED]{quantum electrodynamics}
\acro{QFT}[QFT]{quantum field theory}
\acro{RF}[RF]{radio frequency}
\acro{RPC}[RPC]{pesistive plate chamber}
\acro{SM}[SM]{standard model}
\acro{SPS}[SPS]{Super Proton Synchrotron}
\acro{SUSY}[SUSY]{supersymmetry}
\acro{TEC}[TEC]{tracker endcap}
\acro{TIB}[TIB]{tracker inner barrel}
\acro{TID}[TID]{tracker inner disk}
\acro{TOB}[TOB]{tracker outer barrel}
\acro{UE}[UE]{underlying event}
\acro{VBF}[VBF]{vector boson fusion}
\acro{WLCG}[WLCG]{Worldwide LHC Computing Grid}
\end{acronym}

\appendix
\renewcommand{\chaptername}{Appendix}
\chapter{\texorpdfstring{Systematic uncertainties in the MSSM \AHtotautau analysis}{Overview of systematic uncertainties in the MSSM A/H to tautau analysis}}
\label{appendix:uncerts}
An overview of the systematic uncertainties that are applied in the MSSM \AHtotautau analysis (chapter \ref{chap:mssm})
are summarised in this appendix. The uncertainties are given in
table \ref{tab:SystematicUncertainties_mt} for the $\mu\tau_h$
channel, in table \ref{tab:SystematicUncertainties_et} for the $e\tau_h$ channel,
in table \ref{tab:SystematicUncertainties_tt} for the $\tau_h\tau_h$ channel and
in table \ref{tab:SystematicUncertainties_em} for the $e\mu$ channel. Additionally,
the systematic uncertainties for the $Z\rightarrow\mumu$ control region are shown
in table \ref{tab:SystematicUncertainties_zmm}. Correlations are also indicated.
Some of the uncertainties
are different for different processes, where this is the case ranges are given.
Uncertainties marked "Fully correlated" are correlated between the different
decay channels and event categories, those marked "Fully uncorrelated" are
uncorrelated between the different decay channels and event categories. "Cats:C,chns:U"
indicates that the uncertainty is correlated between categories but uncorrelated between
the different decay channels. For brevity the backgrounds are denoted with shorthand notations.
`ZTT' indicates the \Ztautau component. 'ZJ' is reserved for the \Zellell background
in which the tau is faked by a jet. The \Zellell background in which the tau is faked
by an electron or muon is indicated as 'ZL'. `W' indicates the \Wjets process. `TT' denotes
the \ttbar background, with  `TTT' and `TTJ'
indicating the \ttbar background with real hadronic taus and without real hadronic taus, respectively. 
The di-boson and single-top backgrounds are denoted `VV', with
`VVT' and `VVJ' indicating the di-boson plus single-top background with real hadronic taus
and without real hadronic taus, respectively. The QCD multijet background is denoted `QCD'.
\input{MSSM/UncOv}

%% If you have time and interest to generate a (decent) index,
%% then you've clearly spent more time on the write-up than the 
%% research ;-)
%\printindex