### 2017-06-05 Hhh/Hhhyields.tex

\begin{table}[h!]
\begin{center}
\begin{tabular}{lr@{$ \,\,\pm\,\, $}lr@{$ \,\,\pm\,\, $}lr@{$ \,\,\pm\,\, $}l}
\hline
\multicolumn{7}{c}{\mutau channel} \\
\hline
& \multicolumn{6}{c}{$\sqrt{s}$ = 8~\TeV} \\
Process & \multicolumn{2}{c}{2jet-0tag} & \multicolumn{2}{c}{2jet-1tag} & \multicolumn{2}{c}{2jet-2tag}\\
\hline
QCD                                   & 207     & 33       & 42        & 13 	& 4	& 2 \\
\ttbar                                & 89 	& 9        & 197       & 16     & 109 & 9   \\
Di-bosons + single top                & 38       & 7       & 17        & 2      & 4 &0.6     \\
\Wjets                                & 346     & 32       & 54        & 18     &1 & 7    \\
$\PZ\rightarrow \ell\ell$ (lepton/jet faking $\Pgt$)& 84        & 15        & 6         & 1  &  1 & 0.6       \\
$\PZ\rightarrow \Pgt\Pgt$            & 865       & 56       & 95        & 10  & 12 & 3       \\
\hline
Total Background                      & 1629      & 40       & 410       & 18 & 131 & 14        \\
\hline
\Htohhtobbtautau             & 1        & 0.2        & 5          & 0.3     & 3 & 0.2  \\
\hline
Data                                  & \multicolumn{2}{c}{1622     }& \multicolumn{2}{c}{405      } & \multicolumn{2}{c}{134}\\
\hline
\end{tabular}
\end{center}
\caption{ Number of events observed in data in the 2jet-0tag, 2jet-1tag and 2jet-2tag \mbox{categories} in the \mutau channel, compared
with the background expectation. The signal expectations at \mH~$=300\,\GeV$, \tanb~$=2$ are given for comparison.
The uncertainties
correspond to the post-fit systematic uncerainties, taking into account correlations between them.}
\label{tab:hhh_mt_yields}
\end{table}

\begin{table}[h!]
\begin{center}
\begin{tabular}{lr@{$ \,\,\pm\,\, $}lr@{$ \,\,\pm\,\, $}lr@{$ \,\,\pm\,\, $}l}
\hline
\multicolumn{7}{c}{\etau channel} \\
\hline
& \multicolumn{6}{c}{$\sqrt{s}$ = 8~\TeV} \\
Process & \multicolumn{2}{c}{2jet-0tag} & \multicolumn{2}{c}{2jet-1tag} & \multicolumn{2}{c}{2jet-2tag}\\
\hline
QCD                                   & 219     & 33       & 24        & 8 	& 4	& 5 \\
\ttbar                                & 46 	& 5       & 109       & 9     & 59 & 6   \\
Di-bosons + single top                & 17       & 3       & 8        & 1      & 2 &0.4     \\
\Wjets                                & 186     & 16       & 35       & 13     &3 & 3    \\
$\PZ\rightarrow \ell\ell$ (lepton/jet faking $\Pgt$)& 158        & 26        & 15         & 3  &  1 & 0.5       \\
$\PZ\rightarrow \Pgt\Pgt$            & 393       & 38       & 44        & 5  & 4 & 1       \\
\hline
Total Background                      & 1019      & 31       & 236       & 13 & 73 & 7        \\
\hline
\Htohhtobbtautau             & 1        & 0.1        & 2          & 0.1     & 2 & 0.1  \\
\hline
Data                                  & \multicolumn{2}{c}{1022     }& \multicolumn{2}{c}{237      } & \multicolumn{2}{c}{67}\\
\hline
\end{tabular}
\end{center}
\caption{ Number of events observed in data in the 2jet-0tag, 2jet-1tag and 2jet-2tag \mbox{categories} in the \etau channel, compared
with the background expectation. The signal expectations at \mH~$=300\,\GeV$, \tanb~$=2$ are given for comparison.
The uncertainties
correspond to the post-fit systematic uncerainties, taking into account correlations between them.}
\label{tab:hhh_et_yields}
\end{table}