### 2017-06-05 MSSM/MSSMyields.tex

\begin{table}[h!]
\begin{center}
{\footnotesize
\begin{tabular}{lr@{$ \,\,\pm\,\, $}lr@{$ \,\,\pm\,\, $}l}
\hline
\multicolumn{5}{c}{\mutau channel} \\
\hline
& \multicolumn{4}{c}{$\sqrt{s}$ = 13~\TeV} \\
Process & \multicolumn{2}{c}{No b-tag} & \multicolumn{2}{c}{B-tag}\\
\hline
QCD                                   & 8019       & 492       & 146        & 43        \\
\ttbar (jet faking $\Pgth$)             & 406        & 73        & 83         & 15        \\
\ttbar (genuine $\Pgth$)             & 447        & 21        & 212        & 9         \\
Di-bosons + single top (jet faking $\Pgth$)    & 245        & 13        & 37         & 2         \\
Di-bosons + single top (genuine $\Pgth$)    & 352        & 23        & 63         & 4         \\
\Wjets                            & 8506       & 616       & 116        & 43        \\
$\PZ$+jets (jet faking $\Pgth$)       & 961        & 189       & 12         & 2         \\
$\PZ$+jets (l faking $\Pgth$)         & 4138       & 888       & 41         & 9         \\
$\PZ\rightarrow \Pgt\Pgt$            & 34467      & 896       & 437        & 25        \\
\hline
Total Background                      & 57541      & 301       & 1146       & 36        \\
\hline
A/H$\rightarrow\Pgt\Pgt$             & 259        & 27        & 6          & 0.5       \\
\hline
Data                                  & \multicolumn{2}{c}{57564     }& \multicolumn{2}{c}{1136      }\\
\hline
\end{tabular}
}
\end{center}
\caption{ Number of events observed in data in the b-tag and no b-tag categories in the \mutau
compared with the background expectation. The signal
expectation, for \mA~$=1\,\TeV$ and \tanb~$=50$ in the $m_{\PHiggslight}^{\text{mod}+}$ scenario, is also shown. The uncertainties
correspond to the post-fit systematic uncerainties, taking into account correlations between them.}
\label{tab:mssm_mt_yields}
\end{table}

\begin{table}[h!]
\begin{center}
{\footnotesize
\begin{tabular}{lr@{$ \,\,\pm\,\, $}lr@{$ \,\,\pm\,\, $}l}
\hline
\multicolumn{5}{c}{\etau channel} \\
\hline
& \multicolumn{4}{c}{$\sqrt{s}$ = 13~\TeV} \\
Process & \multicolumn{2}{c}{No b-tag} & \multicolumn{2}{c}{B-tag}\\
\hline
QCD                                   & 5963       & 734       & 82         & 42        \\
\ttbar (jet faking $\Pgth$)             & 262        & 48        & 45         & 8         \\
\ttbar (genuine $\Pgth$)             & 317        & 17        & 150        & 7         \\
Di-bosons + single top (jet faking $\Pgth$)    & 166        & 9         & 24         & 1         \\
Di-bosons + single top (genuine $\Pgth$)    & 238        & 15        & 38         & 2         \\
$\PW$+jets                            & 6778       & 639       & 88         & 47        \\
$\PZ$+jets (jet faking $\Pgth$)       & 734        & 128       & 9          & 2         \\
$\PZ$+jets (l faking $\Pgth$)         & 4517       & 733       & 54         & 8         \\
$\PZ\rightarrow \Pgth\Pgth$            & 12907      & 631       & 174        & 13        \\
\hline
Total Background                      & 31883      & 239       & 664        & 30        \\
\hline
A/H$\rightarrow\Pgt\Pgt$             & 130        & 13        & 4          & 0.3       \\
\hline
Data                                  & \multicolumn{2}{c}{31907     }& \multicolumn{2}{c}{675       }\\
\hline
\end{tabular}
}
\end{center}
\caption{ Number of events observed in data in the b-tag and no b-tag categories in the \etau channel,
compared with the background expectation. The signal
expectation, for \mA~$=1\,\TeV$ and \tanb~$=50$ in the $m_{\PHiggslight}^{\text{mod}+}$ scenario, is also shown. The uncertainties
correspond to the post-fit systematic uncerainties, taking into account correlations between them.}
\label{tab:mssm_et_yields}
\end{table}

\begin{table}[h!]
\begin{center}
{\footnotesize
\begin{tabular}{lr@{$ \,\,\pm\,\, $}lr@{$ \,\,\pm\,\, $}l}
\hline
\multicolumn{5}{c}{\tautau channel} \\
\hline
& \multicolumn{4}{c}{$\sqrt{s}$ = 13~\TeV} \\
Process & \multicolumn{2}{c}{No b-tag} & \multicolumn{2}{c}{B-tag}\\
\hline
QCD multijet                                   & 12489      & 368       & 186        & 18        \\
\ttbar (jet faking $\Pgth$)             & 20         & 3         & 6          & 0.9       \\
\ttbar (genuine $\Pgth$)             & 33         & 3         & 17         & 1         \\
Di-bosons + single top (jet faking $\Pgth$)    & 16         & 0.9       & 3          & 0.2       \\
Di-bosons + single top (genuine $\Pgth$)    & 51         & 5         & 6          & 0.6       \\
$\PW$+jets                            & 409        & 63        & 4          & 0.7       \\
$\PZ$+jets (jet faking $\Pgth$)       & 52         & 9         & 1          & 0.2       \\
$\PZ$+jets (l faking $\Pgth$)         & 124        & 27        & 2          & 0.5       \\
$\PZ\rightarrow \Pgt\Pgt$            & 4127       & 359       & 72         & 7         \\
\hline
Total Background                      & 17321      & 132       & 297        & 16        \\
\hline
A/H$\rightarrow\Pgt\Pgt$             & 107        & 12        & 6          & 0.6       \\
\hline
Data                                  & \multicolumn{2}{c}{17314     }& \multicolumn{2}{c}{298       }\\
\hline
\end{tabular}
}
\end{center}
\caption{ Number of events observed in data in the b-tag and no b-tag categories in the \tautau channel, compared with the background expectation. The signal
expectation, for \mA~$=1\,\TeV$ and \tanb~$=50$ in the $m_{\PHiggslight}^{\text{mod}+}$ scenario, is also shown. The uncertainties
correspond to the post-fit systematic uncerainties, taking into account correlations between them.}
\label{tab:mssm_tt_yields}
\end{table}

\begin{table}[h!]
\begin{center}
{\footnotesize
\begin{tabular}{lr@{$ \,\,\pm\,\, $}lr@{$ \,\,\pm\,\, $}l}
\hline
\multicolumn{5}{c}{\emu channel} \\
\hline
& \multicolumn{4}{c}{$\sqrt{s}$ = 13~\TeV} \\
Process & \multicolumn{2}{c}{No b-tag} & \multicolumn{2}{c}{B-tag}\\
\hline
QCD multijet                                   & 9511       & 653       & 205        & 35        \\
\ttbar                         & 6709       & 210       & 2733       & 65        \\
Di-bosons + single top                & 3798       & 200       & 550        & 30        \\
$\PW$+jets                            & 1675       & 98        & 29         & 2         \\
$\PZ\rightarrow \ell\ell$            & 560        & 32        & 11         & 2         \\
$\PZ\rightarrow \Pgt\Pgt$            & 42463      & 627       & 583        & 29        \\
\hline
Total Background                      & 64717      & 264       & 4112       & 62        \\
\hline
A/H$\rightarrow\Pgt\Pgt$             & 192        & 20        & 5          & 0.3       \\
\hline
Data                                  & \multicolumn{2}{c}{64743     }& \multicolumn{2}{c}{4081      }\\
\hline
\end{tabular}
}
\end{center}
\caption{ Number of events observed in data in the b-tag and no b-tag categories in the and \emu channel, compared with the background expectation. The signal
expectation, for \mA~$=1\,\TeV$ and \tanb~$=50$ in the $m_{\PHiggslight}^{\text{mod}+}$ scenario, is also shown. The uncertainties
correspond to the post-fit systematic uncerainties, taking into account correlations between them.}
\label{tab:mssm_em_yields}
\end{table}